---
title: "Practical Machine Learning Course Project"
author: "Zhiyong Tan"
date: "Monday, December 15, 2014"
output: html_document
---

Download the training and testing data if they weren't already downloaded.
```{r}
if (!file.exists("pml-training.csv")) {
  download.file("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv","pml-training.csv")
  download.file("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv","pml-testing.csv")
}
```

Load the training and testing data.
```{r}
library(dplyr)
training <- tbl_df(read.csv("pml-training.csv"))
testing <- tbl_df(read.csv("pml-testing.csv"))
```

Remove irrelevant data, including indexes (problem_id, X).  
Remove all non-varying data in testing and training datasets.  
Replace invalid values (such as #DIV/0!) with NA.  
Find the columns in the testing data which are all NAs.
```{r}
training[training=="DIV/0!"] <- NA
testing[testing=="DIV/0!"] <- NA

training$X <- NULL
testing$X <- NULL
training$problem_id <- NULL
testing$problem_id <- NULL

# Column indexes that have > 1 unique value in both training and testing sets
# considered separately.
keep <- intersect(as.vector(which(sapply(apply(testing,2,unique),function(x) length(x) > 1))),
                  as.vector(which(sapply(apply(training,2,unique),function(x) length(x) > 1))))
          
# Keep only the aforeementioned columns. Keep classe in training data as well, 
# since it is not in testing data.
training <- training[,c(keep,length(training))]
testing <- testing[,keep]
```

Use a gbm() for the machine learning model, and 5-fold cross-validation to 
estimate out-of-sample error.
```{r}
library(caret)
trainControl <- trainControl(method = "cv",number = 5)
modFit <- train(classe ~ .,method="gbm",data=training,trControl=trainControl)
```

The estimated out-of-sample error is `r signif(1 - max(modFit$results$Accuracy), digits=3)` 
with standard deviation of `r signif(modFit$results$AccuracySD[which.max(modFit$results$Accuracy)], digits=3)`.

The testing predictions are computed as follows:
```{r}
predict(modFit,testing)
```
